---
title: "Community Surveys Visualizations"
author: "Cindy Fajardo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```

#Welcome!
In this repository I will have code to help visualize the raw community survey data collected at Cabrillo National Monument in San Diego, CA. I collected both sessile (percent cover) and mobile species (counts) data in the community found around the limpet population I am following for a year beginning September 2023. 

#Libraries
```{r, Libraries}
library("rerddap")
library(tidyverse)
library("tidync")
library("doParallel")
library(lubridate)
library(patchwork)
library(viridis)
library(here)
library(kableExtra)
library(hrbrthemes)
library(emmeans)
library(agricolae)
library(vegan)
library(performance)
library(broom)
library(car)
library(lme4)
library(lmerTest)
library(pscl)
library(ggridges)
library(devtools)
library(pairwiseAdonis)
```

#Read in CSV
```{r, Read in CSV}
feb_cnm_sessile <- read_csv(here("Data", "2024_02_CNM_Sessile_Data.csv"))
feb_cnm_mobiles <- read_csv(here("Data","2024_02_CNM_Mobile_Data.csv"))
feb_cnm_mob_count_mat <- read_csv(here("Data", "2024_02_CNM_Mobile_Count_Matrix.csv"))
feb_cnm_sessile_count_mat <- read_csv(here("Data", "2024_02_CNM_Sessile_Count_Matrix.csv"))
sessile_master <- read_csv(here("Data", "2023_09_CNM_Sessile_Data.csv"))
sess_percentcover <- read_csv(here("Data", "Master_CNM_Sessile_Percent_Data.csv"))
mobiles_master <- read.csv(here("Data", "Master_CNM_Mobile_Data.csv"))
```

#Data Manipulation of Mobile Species
In this chunk, I manipulate the data to help visualize it
```{r, #Data Manipulation}
feb_cnm_mobiles_sum_spec <- feb_cnm_mobiles %>% 
  mutate(spec_sum = rowSums(across(c(M_lugubris, L_dentiens, L_plena, L_keenae, L_limatula, L_scabra, Nuttallina_sp, Mopalia_sp, P_crassipes, T_eiseni, T_funebralis, T_gallina, L_gigantea)), na.rm=TRUE)) %>% 
   pivot_longer(cols = M_lugubris:L_gigantea, #pivoted to make longer
               names_to = "Species", #all species name in this column
               values_to = "Count")
```


#Boxplot of Mobile Organism Abundance Per Plot in Presence/Absence of Central Limpet
```{r,Boxplot of Mobile Organism Abundance Per Plot in Presence/Absence of Central Limpet}
 feb_cnm_mobiles_sum_spec %>% 
  ggplot(aes(
    x = COL_pres,
    y = spec_sum,
    fill = COL_pres))+
  geom_boxplot(show.legend = FALSE)+ 
  scale_fill_brewer(palette = "Dark2")+
  labs(x = "Presence of Central Owl Limpet") +
  theme_bw()
  
```

#Data Manipulation of Sessile Species
Here, I manipulate the data to visualize the data. I separate it by animals and algae. 
```{r, #Data Manipulation of Sessile Species}
feb_cnm_sessile_sum_plot <- feb_cnm_sessile %>%
  mutate(spec_sum = rowSums(across(c(A_elegantissima, Phragmatopoma_californica, Balanus_Chtalmus, Bare_rock, Ralfsia_sp, Coralline_crust, Endocladia_sp, P_polymerus,Ulva_sp)), na.rm=TRUE))%>% 
  mutate(Animal = (A_elegantissima+ Phragmatopoma_californica+ Balanus_Chtalmus+ P_polymerus)/spec_sum*100,
         Algae = (Coralline_crust+Endocladia_sp+Ulva_sp)/spec_sum*100)
```

#Boxplot of Sessile Organism Percent Cover Per Plot in Presence/Absence of Central Limpet
```{r, Boxplot of Sessile Organism Percent Cover Per Plot in Presence/Absence of Central Limpet}
feb_cnm_sessile_sum_plot %>% 
  pivot_longer(cols = c(Animal, Algae), names_to = "Percent_Type")%>%
  ggplot(aes(
    x = COL_pres,
    y = value,
    fill = Percent_Type))+
  geom_boxplot(show.legend = TRUE)+ 
  scale_fill_brewer(palette = "Pastel2")+
  scale_color_brewer(palette = "Pastel1")+
  labs(x = "Presence of Central Owl Limpet",
       fill = "Percent Type") +
  theme_bw()

```

#Data Manipulation for Species Richness
```{r, Data Manipulation for Species Richness}
feb_cnm_mob_count_mat$COL_pres <- as.factor(feb_cnm_mob_count_mat$COL_pres) #made this coulmn a factor
feb_cnm_mob_count_mat$Transect<- as.factor(feb_cnm_mob_count_mat$Transect)

#USE THIS ONE
feb_cnm_mob_count_mat_sum <- feb_cnm_mob_count_mat %>% 
  mutate(spec_sum = rowSums(across(c(M_lugubris, L_dentiens, L_plena, L_keenae, L_limatula, L_scabra, Nuttallina_sp, Mopalia_sp, P_crassipes, T_eiseni, T_funebralis, T_gallina, L_gigantea)), na.rm=TRUE))

```

#Species Richness of Mobiles Boxplot
```{r, Species Richness of Mobiles Boxplot}
feb_mobile_spec_rich_box <- feb_cnm_mob_count_mat_sum %>% 
  ggplot(aes(
    x = COL_pres,
    y = spec_sum,
    fill = COL_pres))+
  geom_boxplot(show.legend = FALSE)+ 
  scale_fill_brewer(palette = "Dark2")+
  labs(x = "Presence")+
  theme_bw()
feb_mobile_spec_rich_box
```

#Sessile Species Richnnes Data Manipulation 
```{r, Sessile Species Richnnes Data Manipulation}
feb_cnm_sessile_count_mat$COL_pres <- as.factor(feb_cnm_sessile_count_mat$COL_pres) #made this coulmn a factor
feb_cnm_sessile_count_mat$Transect<- as.factor(feb_cnm_sessile_count_mat$Transect)

#Using this one
feb_cnm_sessile_count_mat_sum <- feb_cnm_sessile_count_mat %>% 
    select(Transect, Plot_Name, L_R, Meter_no, COL_pres, Tag_ID, Anthopleura_elegantissima,Phragmatopoma_californica, Balanus_glandula, Chthalmus, Brown_encrusting, Coralline_crust, Endocladia_sp, Goose_neck, Ulva ) %>% 
  mutate(spec_sum = rowSums(across(c(Anthopleura_elegantissima,Phragmatopoma_californica, Balanus_glandula, Chthalmus, Brown_encrusting, Coralline_crust, Endocladia_sp, Goose_neck, Ulva)), na.rm=TRUE))

####Try this###
feb_cnm_sessile_matrix <- feb_cnm_sessile_count_mat %>%
  mutate(spec_sum = rowSums(across(c(Anthopleura_elegantissima,Phragmatopoma_californica, Balanus_glandula, Chthalmus, Bare_rock, Brown_encrusting, Coralline_crust, Endocladia_sp, Goose_neck, Ulva)), na.rm=TRUE))%>% 
  mutate(Animal = (Anthopleura_elegantissima+ Phragmatopoma_californica+ Balanus_glandula+ Goose_neck+ Chthalmus),
         Algae = (Brown_encrusting+Coralline_crust+Endocladia_sp+Ulva))

```

#Boxplot of Sessile Species Richness
```{r, Boxplot of Sessile Species Richness}
feb_cnm_sessile_matrix %>% 
  pivot_longer(cols = c(Animal, Algae), names_to = "sessile_type")%>%
  ggplot(aes(
    x = COL_pres,
    y = value,
    fill = sessile_type))+
  geom_boxplot(show.legend = TRUE)+ 
  scale_fill_brewer(palette = "Pastel1")+
  scale_color_brewer(palette = "Set1")+
  labs(x = "Presence of Central Owl Limpet",
       fill = "Percent Type") +
  theme_bw()

```

#Patchwork of Mobile Abundance and Sessile Percent Cover
```{r, Patchwork of Mobile Abundance and Sessile Percent Cover}
 patch_mob <- feb_cnm_mobiles_sum_spec %>% 
  ggplot(aes(
    x = COL_pres,
    y = spec_sum,
    fill = COL_pres))+
  geom_boxplot(show.legend = FALSE)+ 
  scale_fill_brewer(palette = "Dark2")+
  labs(title = "Abundance of Mobile Species",
       x = "Presence of Central Owl Limpet",
       y = "Abundance of Other Mobile Organisms") +
  theme_bw()

patch_sess <- feb_cnm_sessile_sum_plot %>% 
  pivot_longer(cols = c(Animal, Algae), names_to = "percent_type")%>%
  ggplot(aes(
    x = COL_pres,
    y = value,
    fill = COL_pres,
    color = percent_type))+
  geom_boxplot(show.legend = TRUE)+ 
  scale_fill_brewer(palette = "Pastel1")+
  scale_color_brewer(palette = "Set1")+
  labs(title = "Percent Cover of Sessile Species",
       x = "Presence of Central Owl Limpet",
       y = "Percent Cover") +
  theme_bw()

patch_mob+patch_sess+
  plot_annotation(title = 'Winter Total Aundance of Mobiles Species and Percent Cover of Sessile Species in the Presence/Absence of Owl Limpets', theme = theme(plot.title = element_text(size = 12)))
```

#Patchwork of species richness for mobiles and sessile
```{r, Patchwork of species richness for mobiles and sessile}
mobile_spec_rich_box_nolab <- feb_cnm_mob_count_mat_sum %>% 
  ggplot(aes(
    x = COL_pres,
    y = spec_sum,
    fill = COL_pres))+
  geom_boxplot(show.legend = FALSE)+ 
  scale_fill_brewer(palette = "Dark2")+
  labs(title = "Mobiles",
       x = "Presence/Absence",
       y = "Species Richness")+
  theme_bw()

sessile_spec_rich_nolab <- feb_cnm_sessile_count_mat_sum %>% 
  ggplot(aes(
    x = COL_pres,
    y = spec_sum,
    fill = COL_pres))+
  geom_boxplot(show.legend = FALSE)+
  scale_fill_brewer(palette = "Dark2")+
  labs(title = "Sessile",
       x = "Presence/Absence",
       y = "Species Richness")+
  theme_bw()


mobile_spec_rich_box_nolab+sessile_spec_rich_nolab+
  plot_annotation(title = 'Winter Species Richness of Mobiles and Sessile Species in the Presence/Absence of Owl Limpets',
                  theme = theme(plot.title = element_text(size = 15)))
```
